DS-GA 1013: Project\\Intelligible Principal Component Analysis

### 2020-05-08 Paper/Paper.tex
Updates from Overleaf

\documentclass[11pt,letterpaper]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{upgreek}
\usepackage{apacite}
\usepackage{booktabs}
\usepackage{ntheorem}
\usepackage{enumitem}
\usepackage{gensymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\graphicspath{ {./Figures/} }
\newtheorem{hyp}{Hypothesis}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{
	Kho, Lee\\
	\textit{ltk224}
	\and
	Peeters, Jonas\\
	\textit{jp5642}
}
\title{DS-GA 1013: Project\\Intelligible Principal Component Analysis}

@@

@@

\textit{Total Explained Variance:} Most of the method we review in this paper yield orthogonal principal components. Thus, they will span the observation space of $X$ explaining all the available variance among themselves. However, the SPCA method does enforce orthogonality upon its principal components and hence these sparse components will not explained all the variance in our data set. We calculate the total variance explained by the non-orthogonal components as the total variance explained by the orthogonal basis of the span of these non-orthogonal components. This orthogonal basis is derived using SVD.\\

@@

The table below gives a summary of the results of the different sparse PCA methods for varying regularization parameters.\\

{\color{red}[TABLE HERE]
We want four rows: PCA, SCoT, SCoTLASS, SPCA for their "best" specification. We will show the difference within each model in tables and figures in the appendix. Columns containing: Weighted Gini coefficient, distribution of explained variance, }
\\

\begin{table}[htpb!]
    \centering
    \caption{Variables ordered by loading magnitude (1 = largest loading weight)}
    \label{rstl:loadings}
    \input{../Tables/orderedLoads.tex}
\end{table}

@@

\section*{Discussion}
%Discussion: What did you find out? Do the results suggest any other interesting questions to explore?

@@

>>>>

\documentclass[11pt,letterpaper]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{upgreek}
\usepackage{apacite}
\usepackage{booktabs}
\usepackage{ntheorem}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{gensymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\graphicspath{ {./Figures/} }
\newtheorem{hyp}{Hypothesis}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{
	Kho, Lee\\
	\textit{ltk224}
	\and
	Peeters, Jonas\\
	\textit{jp5642}
}
\title{DS-GA 1013: Project\\Intelligible Principal Component Analysis}

@@

\makeatletter
\def\@makechapterhead#1{%
  %%%%\vspace*{50\p@}% %%% removed!
  {\parindent \z@ \raggedright \normalfont
    \ifnum \c@secnumdepth >\m@ne
        \huge\bfseries \@chapapp\space \thechapter
        \par\nobreak
        \vskip 20\p@
    \fi
    \interlinepenalty\@M
    \Huge \bfseries #1\par\nobreak
    \vskip 40\p@
  }}
\def\@makeschapterhead#1{%
  %%%%%\vspace*{50\p@}% %%% removed!
  {\parindent \z@ \raggedright
    \normalfont
    \interlinepenalty\@M
    \Huge \bfseries  #1\par\nobreak
    \vskip 40\p@
  }}
\makeatother

@@

\textit{Total Explained Variance:} Most of the method we review in this paper yield orthogonal PCs. Thus, summing the explained variance of each resulting PC gives us the total variance of the data. However, the SPCA method does not enforce orthogonality upon its PCs and hence summing the explained variance of each resulting PC would be too generous an estimate for total variance. Therefore, we calculated \textit{adjusted explained variance} for the each method, as described in \citeyear{Shen2008}, which reflects the incremental increase in explained variance that each additional PC provides (see Appendix for adjusted explained variance algorithm). For SCoT and SCoTLASS, adjusted explained variance is the same as regular explained variance.\\

@@

Table \ref{tab:sumStats} shows a summary of the results of the different sparse PCA methods for varying regularization parameters.\\

\begin{table}[htpb!]
    \begin{minipage}{.49\linewidth}
        \caption{Summary Statistics}
        \label{tab:sumStats}
        \centering
        \input{../Tables/SummaryModels.tex}
    \end{minipage}
    \begin{minipage}{.49\linewidth}
        \caption{Sorted Loadings}
        \label{tab:sorLoad}
        \centering
        \input{../Tables/orderedLoads.tex}
    \end{minipage}
\end{table}

@@

\begin{figure}[htpb!]\centering
    \caption{Sparsity of Components}
    \label{fig:sparComp}
    \includegraphics[scale = 0.5]{"../Figures/sparsityOfComponents.pdf"}
\end{figure}

\section*{Discussion}
We notice while PCA delivers us orthogonal principal components with optimal front-loaded explained variance, they are unintelligible. Over the years several sparse PCA algorithms have been introduced such as SCoT, SCoTLASS and SPCA. We find that all of them make a sparsity versus front-loaded variance trade-off. After comparing all models we conclude that SPCA comes out as the first his class with {\color{red} continue...}, but does not yield orthogonal components like the other options.\\

Since SPCA's components are correlated, an interesting future topic of research could be finding a way to transform these into orthogonal set of SPCA components. Moreover, while LARS and proximal gradient descent can be used to solve for a LASSO component, a lot of efficiency could be gained by a faster algorithm. As we tried working with a dataset of $>20k$ observations, we notice the many limitations these wide dataset enforce on PCA methods and their computability.
%Discussion: What did you find out? Do the results suggest any other interesting questions to explore?

@@

\begin{figure}
\begin{subfigure}[htpb!]{0.49\textwidth}
    \centering
    \caption{Explained Variance for SCoT}
    \label{fig:EV_SCoT}
    \includegraphics[scale = 0.5]{"../Figures/SCoT_comparison.pdf"}
\end{subfigure}
\begin{subfigure}[htpb!]{0.49\textwidth}
    \centering
    \caption{Explained Variance for SCoTLASS}
    \label{fig:EV_SCoTLASS}
    \includegraphics[scale = 0.5]{"../Figures/SCoTLASS_comparison.pdf"}
\end{subfigure}
\end{figure}
